---
title: "math regression"
output: pdf_notebook
---
author: "hz2482"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load library
```{r}
library('rminer')
library('ModelMetrics')
```
# set up output matrix and cross-validation parameter
```{r parameter}
# 10-fold cross-validation
K=c('kfold',10)
# output matrix
reg.result=matrix(nrow=3,ncol=5)
reg.t=matrix(nrow=3,ncol=5)
colnames(reg.result)=c('NV','NN','SVM','DT','RF')
rownames(reg.result)=c('A','B','C')
colnames(reg.t)=c('NV','NN','SVM','DT','RF')
rownames(reg.t)=c('A','B','C')
```
# A input setup
# A-naive predictor
```{r A-NV}
# G3^=G2
reg.result['A','NV']=rmse(data_math$G3,data_math$G2)
A.NV=mining(G3~.-bin-l5,data_math,model='naive',Runs=20,task='reg',method=K)
# rmse and t-test
mean(mmetric(A.NV,metric=c("RMSE"))$RMSE1)
reg.t['A','NV']=0
```
# A-neural networks
```{r A-NN}
# neural networks with 100 epochs, one layer and 0, 2, 4, 6, 8 nodes
A.s_nn=list(smethod='grid',search=list(size=seq(0,8,2)),method=K,convex=0)
A.NN=mining(G3~.-bin-l5,data_math_2,model='mlp',search=A.s_nn,maxit=100,scale='input',task='reg',Runs=20,method=K)
savemining(A.NN,'ANN',ascii=TRUE)
# rmse and t-test
A.m_nn=mmetric(A.NN,metric=c("RMSE"))
reg.result['A','NN']=mean(A.m_nn$RMSE1)
reg.t['A','NN']=abs(t.test(A.m_nn$RMSE1)$conf.int[1]-mean(A.m_nn$RMSE1))
```
# A-support vector machine
```{r A-SVM}
# support vector machine with gaussian kernel and hyper paremeter -9,-7,-5,-3,-1
A.s_svm=list(smethod='grid',search=list(sigma=2^c(-9,-7,-5,-3,-1)),method=K,convex=0)
A.SVM=mining(G3~.-bin-l5,data_math_2,model='ksvm',kernel='rbfdot',search=A.s_svm,scale='input',task='reg',Runs=20,method=K)
savemining(A.SVM,'ASVM',ascii=TRUE)
# rmse and t-test
A.m_svm=mmetric(A.SVM,metric=c("RMSE"))
reg.result['A','SVM']=mean(A.m_svm$RMSE1)
reg.t['A','SVM']=abs(t.test(A.m_svm$RMSE1)$conf.int[1]-mean(A.m_svm$RMSE1))
```
# A-decision tree
```{r A-dt}
# decision tree
A.DT=mining(G3~.-bin-l5,data_math,model='dt',Runs=20,task='reg',method=K)
savemining(A.DT,'ADT',ascii=TRUE)
# rmse and t-test
A.m_dt=mmetric(A.DT,metric=c("RMSE"))
reg.result['A','DT']=mean(A.m_dt$RMSE1)
reg.t['A','DT']=abs(t.test(A.m_dt$RMSE1)$conf.int[1]-mean(A.m_dt$RMSE1))
```
# A-random forest
```{r A-rf}
# random forest
A.RF=mining(G3~.-bin-l5,data_math,model='randomforest',Runs=20,task='reg',method=K)
savemining(A.RF,'ARF',ascii=TRUE)
# rmse and t-test
A.m_rf=mmetric(A.RF,metric=c("RMSE"))
reg.result['A','RF']=mean(A.m_rf$RMSE1)
reg.t['A','RF']=abs(t.test(A.m_rf$RMSE1)$conf.int[1]-mean(A.m_rf$RMSE1))
```
# B input setup
# B-naive predictor
```{r B-NV}
# G3^=G1
reg.result['B','NV']=rmse(data_math$G3,data_math$G1)
# rmse and t-test
reg.t['B','NV']=0
```
# B-neural networks
```{r B-NN}
# neural networks with 100 epochs, one layer and 0, 2, 4, 6, 8 nodes
B.s_nn=list(smethod='grid',search=list(size=seq(0,8,2)),method=K,convex=0)
B.NN=mining(G3~.-bin-l5-G2,data_math_2,model='mlp',search=B.s_nn,maxit=100,scale='input',task='reg',Runs=20,method=K)
savemining(B.NN,'BNN',ascii=TRUE)
# rmse and t-test
B.m_nn=mmetric(B.NN,metric=c("RMSE"))
reg.result['B','NN']=mean(B.m_nn$RMSE1)
reg.t['B','NN']=abs(t.test(B.m_nn$RMSE1)$conf.int[1]-mean(B.m_nn$RMSE1))
```
# B-support vector machine
```{r B-SVM}
# support vector machine with gaussian kernel and hyper paremeter -9,-7,-5,-3,-1
B.s_svm=list(smethod='grid',search=list(sigma=2^c(-9,-7,-5,-3,-1)),method=K,convex=0)
B.SVM=mining(G3~.-bin-l5-G2,data_math_2,model='ksvm',kernel='rbfdot',search=B.s_svm,scale='input',task='reg',Runs=20,method=K)
savemining(B.SVM,'BSVM',ascii=TRUE)
# rmse and t-test
B.m_svm=mmetric(B.SVM,metric=c("RMSE"))
reg.result['B','SVM']=mean(B.m_svm$RMSE1)
reg.t['B','SVM']=abs(t.test(B.m_svm$RMSE1)$conf.int[1]-mean(B.m_svm$RMSE1))
```
# B-decision tree
```{r B-dt}
# decision tree
B.DT=mining(G3~.-bin-l5-G2,data_math,model='dt',Runs=20,task='reg',method=K)
savemining(B.DT,'BDT',ascii=TRUE)
# rmse and t-test
B.m_dt=mmetric(B.DT,metric=c("RMSE"))
reg.result['B','DT']=mean(B.m_dt$RMSE1)
reg.t['B','DT']=abs(t.test(B.m_svm$RMSE1)$conf.int[1]-mean(B.m_svm$RMSE1))
```
# B-random forest
```{r B-rf}
# random forest
B.RF=mining(G3~.-bin-l5-G2,data_math,model='randomforest',Runs=20,task='reg',method=K)
savemining(B.RF,'BRF',ascii=TRUE)
# rmse and t-test
B.m_rf=mmetric(B.RF,metric=c("RMSE"))
reg.result['B','RF']=mean(B.m_rf$RMSE1)
reg.t['B','RF']=abs(t.test(B.m_svm$RMSE1)$conf.int[1]-mean(B.m_svm$RMSE1))
```
# C input setup
# C-naive predictor
```{r C-NV}
# G3^=average(G3)
reg.result['C','NV']=rmse(data_math$G3,rep(mean(data_math$G3),395))
C.NV=mining(G3~.-bin-l5-G2-G1,data_math,model='naive',Runs=20,task='reg',method=K)
# rmse and t-test
mean(mmetric(C.NV,metric=c("RMSE"))$RMSE1)
reg.t['C','NV']=0
```
# C-neural networks
```{r C-NN}
# neural networks with 100 epochs, one layer and 0, 2, 4, 6, 8 nodes
C.s_nn=list(smethod='grid',search=list(size=seq(0,8,2)),method=K,convex=0)
C.NN=mining(G3~.-bin-l5-G2-G1,data_math_2,model='mlp',search=C.s_nn,maxit=100,scale='input',task='reg',Runs=20,method=K)
savemining(C.NN,'CNN',ascii=TRUE)
# rmse and t-test
C.m_nn=mmetric(C.NN,metric=c("RMSE"))
reg.result['C','NN']=mean(C.m_nn$RMSE1)
reg.t['C','NN']=abs(t.test(C.m_nn$RMSE1)$conf.int[1]-mean(C.m_nn$RMSE1))
```
# C-support vector machine
```{r C-SVM}
# support vector machine with gaussian kernel and hyper paremeter -9,-7,-5,-3,-1
C.s_svm=list(smethod='grid',search=list(sigma=2^c(-9,-7,-5,-3,-1)),method=K,convex=0)
C.SVM=mining(G3~.-bin-l5-G2-G1,data_math_2,model='ksvm',kernel='rbfdot',search=C.s_svm,scale='input',task='reg',Runs=20,method=K)
savemining(C.SVM,'CSVM',ascii=TRUE)
# rmse and t-test
C.m_svm=mmetric(C.SVM,metric=c("RMSE"))
reg.result['C','SVM']=mean(C.m_svm$RMSE1)
reg.t['C','SVM']=abs(t.test(C.m_svm$RMSE1)$conf.int[1]-mean(C.m_svm$RMSE1))
```
# C-decision tree
```{r C-dt}
# decision tree
C.DT=mining(G3~.-bin-l5-G2-G1,data_math,model='dt',Runs=20,task='reg',method=K)
savemining(C.DT,'CDT',ascii=TRUE)
# rmse and t-test
C.m_dt=mmetric(C.DT,metric=c("RMSE"))
reg.result['C','DT']=mean(C.m_dt$RMSE1)
reg.t['C','DT']=abs(t.test(C.m_dt$RMSE1)$conf.int[1]-mean(C.m_dt$RMSE1))
```
# C-random forest
```{r C-rf}
# random forest
C.RF=mining(G3~.-bin-l5-G2-G1,data_math,model='randomforest',Runs=20,task='reg',method=K)
savemining(C.RF,'CRF',ascii=TRUE)
# rmse and t-test
C.m_rf=mmetric(C.RF,metric=c("RMSE"))
reg.result['C','RF']=mean(C.m_rf$RMSE1)
reg.t['C','RF']=abs(t.test(C.m_rf$RMSE1)$conf.int[1]-mean(C.m_rf$RMSE1))
```
# save result
```{r save_result}
write.csv(reg.result,"reg.csv")
write.csv(reg.t,"reg_t.csv")
```

